---
title: "Lab5"
author: "Astros"
date: "12/15/2020"
output: 
  html_document:
    keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE,warning=FALSE)

```



```{r}

library(dplyr)

```


###Data import and aggregation


```{r}

inv.dat <- read.csv("data.investors16.csv", header = F, row.names = 1)
dim(inv.dat) #check dimensions

xbarA <- apply(inv.dat, MARGIN = 1, mean) #sample mean of A for each day
ssdA <- apply(inv.dat, MARGIN = 1, sd) #std dev of A for each day
nobsA <- apply(inv.dat, MARGIN = 1, length) #sample size of A for each day

mystatsA <- data.frame(xbarA, ssdA, nobsA)


inv.dat.tall <- read.csv("data.investors.tall.csv", header = T)
dim(inv.dat.tall) #check dimensions


xbarB <- inv.dat.tall %>% group_by(day) %>% summarise(mean = mean(x))
ssdB <- inv.dat.tall %>% group_by(day) %>% summarise(sd = sd(x))
nobsB <- inv.dat.tall %>% group_by(day) %>% summarise(count = length(x))

mystatsB <- data.frame(xbarB = xbarB$mean, ssdB = ssdB$sd,nobsB = nobsB$count, row.names = xbarB$day)
row.names(mystatsB) <- row.names(mystatsA)


#Confirm that mystatsA matches mystatsB

cbind(mystatsA,mystatsB)

TF <- mystatsA == mystatsB 
TF

#create temporary data frames to make the dimnames equal to one another without affecting code further down; this is just to test that the values are the same using identical() function
tmp_mystatsA <- mystatsA
tmp_mystatsB <- mystatsB
dimnames(tmp_mystatsB) <- dimnames(tmp_mystatsA)
identical(tmp_mystatsA, tmp_mystatsB) 


# another method to show both the data frames are equal

identical(mystatsA$xbarA,mystatsB$xbarB)
identical(mystatsA$ssdA,mystatsB$ssdB)
identical(mystatsA$nobsA,mystatsB$nobsB)


```


###Investor A
Run mystats = mystatsA to make a new data.frame. (Would it matter if you used mystatsB instead?)
We will use mystats for everything that follows. Each step below asks you to compute a 20-vector and
add it to mystats. The column dimension of mystats will go from 3 to 10.



```{r}

mystats = mystatsA
mystats

```


Compute the value of the test statistic for each of the 20 days.


```{r}

mystats <- mystats %>% mutate(tstat=(xbarA-5)/(ssdA/sqrt(nobsA)))
row.names(mystats) <- row.names(mystatsA)
mystats

```


Let tcrit1 be a 20-vector containing
the negative critical value, and let tcrit2 contain the positive critical value. Add these two vectors to
mystats.


```{r}

alpha <- 0.10
df <- nobsA-1
mystats <- mystats %>% mutate(tcrit1=qt(alpha/2,df))
mystats <- mystats %>% mutate(tcrit2=qt(alpha/2,df,lower.tail=F))
row.names(mystats) <- row.names(mystatsA)
mystats


```


Perform the hypothesis test. If tstat is outside the critical values, Investor A rejects H0. If tstat is
between the critical values, she accepts. Use this rule with ifelse() to create vector test1 populated
with the values ACCEPT or REJECT.


```{r}

mystats$test1 <- ifelse((mystats$tstat > mystats$tcrit1) & (mystats$tstat < mystats$tcrit2), "ACCEPT", "REJECT") 
mystats

```


The p-value represents the probability of obtaining getting a value of the test statistic more contradictory
to H0 than tstat. Compute the vector of p-values like this:


```{r}

mystats <- mystats %>% mutate(pval=2*(1-pt(abs(mystats$tstat), nobsA - 1)))
row.names(mystats) <- row.names(mystatsA)
mystats


```


Use the p-value to perform the hypothesis test a different way. Compare the p-value to the significance
level α. If pval ≤ α, reject H0. Otherwise, accept H0. Use this rule with ifelse() to create vector
test2 populated with the values ACCEPT or REJECT. It should be identical to test1.


```{r}

mystats$test2 <- ifelse(mystats$pval <= 0.1, "REJECT", "ACCEPT")   #perform the hypothethis test using p-value
mystats 

```


 Finally, create vector dec.A and populate it with the values GOLF or STARBUCKS. The data.frame
mystats should now be 20x10.


```{r}

mystats$dec.A <- ifelse(mystats$test1 == "ACCEPT", "GOLF", "STARBUCKS")
mystats

```



Investor B


```{r}


df <- nobsA-1

(t <- round(qt((1-0.10/2),df),2))



(mystats$lim1 <- mystatsB$xbarB - t*(mystatsB$ssdB/(sqrt(mystatsB$nobsB))))

(mystats$lim2 <- mystatsB$xbarB + t*(mystatsB$ssdB/(sqrt(mystatsB$nobsB))))



mystats$test3 <- ifelse(mystats$lim2 < 5, "REJECT", "ACCEPT")

mystats$dec.B <- ifelse(mystats$test3 == "ACCEPT", "GOLF", "STARBUCKS")

mystats


```



####A level questions


Do you think A and B will ever get to know each other? Explain why or why not

```{r}

c(paste("The number of days that investigator A and B are in the same place out of 20 days is :" ))

sum(mystats$dec.A == mystats$dec.B)

  

```


Assume the true value of µ is 5 on each of the 20 trading days. Use that information to add columns
grade.A and grade.B to mystats. Grade each investor’s decisions as CORRECT or INCORRECT.

```{r}


mystats$grade.A <- ifelse(mystats$test1 == "REJECT", "INCORRECT", "CORRECT") 

mystats$grade.B <- ifelse(5 > mystats$lim2 | 5 < mystats$lim1, "INCORRECT", "CORRECT")

mystats

```


Suppose Investor B uses α = .05 on each of the 20 days. Run his analyses again with this new α value
and make a new version of mystats to incorporate the new results


```{r}

#alpha = 0.05
df <- nobsA-1

(t1 <- round(qt((1-0.05/2),df),2))

mystats$newlim1 <- mystatsB$xbarB - t1*(mystatsB$ssdB/(sqrt(mystatsB$nobsB)))

mystats$newlim2 <- mystatsB$xbarB + t1*(mystatsB$ssdB/(sqrt(mystatsB$nobsB)))

mystats$newtest <- ifelse(mystats$newlim2 >= 5, "ACCEPT", "REJECT")
mystats$dec.BNew <- ifelse(mystats$newtest == "ACCEPT", "GOLF", "STARBUCKS")
mystats$grade.BNew <- ifelse(5 > mystats$newlim2 | 5 < mystats$newlim1, "INCORRECT", "CORRECT")

c(paste("The chances of A and B getting to know each other is the same as before." ))

sum(mystats$dec.A == mystats$dec.BNew)


mystats


```





